---
title: "Summary of Scope 1 and 2 Emissions from EAF Steel Plants, 2021"
author: "Ruth Enriquez, Erica Bishop, Michael Zargari, and Amrit Sandhu"
format: pdf
editor: visual
---

```{r loading_packages}
library(tidyverse)
library(here)
library(ggpubr)
library(janitor)
```

```{r read in output data}

#loading data outputs steel_co2e_2021_analysis
annual_data <- read_csv(here("..", "output_data", "2021_steel_plant_emissions_and_productions.csv"))

```
### Summary by Plant

Show the top 5 biggest emitting plants
Show top 5 highest emissiosn intensity plants
Show the average scope 1 v. scope 2 ratio for plant emissiosn of total

```{r}

#top 5 highest emitting plants by total emissions
top_plants <- annual_data %>% 
  group_by(plant_name) %>% 
  summarise(total)




```


### Summary by Company

-   Table of summary statistics: average and median Scope 1, Scope 2, and total emissions for each company

```{r get company }

# creating summary statistic table
top5_companies <- annual_data %>% 
  group_by(owner) %>% 
  summarise(
    scope_2_emissions = max(scope2_co2e_tonnes_2021),
    scope_1_emissions = max
  )


# owner_stats <- annual_data_full %>% 
#          group_by(owner) %>% 
#          summarise(co2_intensity_avg = mean(emissions_intensity_co2e_tonne_per_tonne),
#                    allplants_scope1_and2 = sum(total_co2e_tonnes_2021),
#                    allplants_scope2_total = sum(scope2_co2e_tonnes_2021),
#                    allplants_scope1_total = sum(scope1_co2e_tonnes_2021))

```

-   Top 5 emitting companies

-   Benchmarking companies

-   Comparison table of companies (all)

### Summary by Region

-   Emissions intensities by region 
- emissions by region

```{r regions}

regional_emissions <- annual_data %>% 
  group_by(region)


```


### Conclusion:

Call for data transparency for the steel industry and beyond. Electrification doesn't equal decarbonization - we need the data to prove it.