---
title: "state_emissions_per_plant"
format: html
editor: visual
---

### Loading Libraries

```{r, message=FALSE}
library(terra)
library(sf)
library(raster)
library(here)
library(janitor)
library(tidyverse)
library(tmap)
library(rmapshaper)
```

### Importing eGrid Data

```{r}
## Importing grid region electricity values
egrid2021_data <- readxl::read_excel(here("data", "eGRID2021_data.xlsx"), sheet = "SRL21") %>%
  clean_names()
```

### Standardizing state names and cleaning up rows

```{r}
## Editing egrid data to be how we would like to use it
emissions_by_subregion <- egrid2021_data %>%
  mutate(co2e_lbs_per_mwh = as.numeric(e_grid_subregion_annual_co2_equivalent_combustion_output_emission_rate_lb_m_wh)
         , subregion = e_grid_subregion_acronym
         , subregion_name = e_grid_subregion_name
         , co2e_tonnes_per_mwh = co2e_lbs_per_mwh / 2204.62262185) %>%
  filter(!subregion %in% c("AKGD", "AKMS", "HIMS", "HIOA", "PRMS")) %>% 
  select(c(subregion, subregion_name, co2e_tonnes_per_mwh))

## Removing the first row since it is a part of the column names
emissions_by_subregion <- emissions_by_subregion[-1,]
```

### Importing GEM data

```{r}
gem_data_readin <- readxl::read_excel(here("data", "GEM_2022_data.xlsx"), sheet = "Steel Plants") %>% 
  clean_names()
```

### Filtering to only look at plants and data that we are interested in

```{r}
gem_data <- gem_data_readin %>% 
  filter(country == "United States"
         , status == "operating"
         , nominal_eaf_steel_capacity_ttpa != "N/A") %>% 
  mutate(plant_name = plant_name_english
         , state = subnational_unit_province_state
         , eaf_capacity = as.numeric(nominal_eaf_steel_capacity_ttpa)
         , address = location_address) %>%
  select(plant_id
         , plant_name
         , owner
         , coordinates
         , country
         , state
         , status
         , start_date
         , plant_age_years
         , eaf_capacity
         , municipality
         , address)
```

### Only selecting columns we need for now

```{r}
plant_capacity <- gem_data %>% 
  select(-c(country
         , owner
         , start_date
         , plant_age_years
         , status)) %>% 
  separate(coordinates, c("lat", "lon"), remove = TRUE, ",")
```

### Finding where plants are located in subregions

```{r}
## Reading in data
subregion_shapes_raw <- read_sf(here("..", "data", "egrid2020_subregions", "eGRID2020_subregions.shp")) %>% 
  clean_names() %>% 
  filter(!zip_subregi %in% c("AKGD", "AKMS", "HIMS", "HIOA", "PRMS"))

## Making our file easier to load and map
subregion_shapes <- ms_simplify(subregion_shapes_raw, keep = 0.0005)

## Bringing in our points
plant_points <- st_geometry(st_as_sf(plant_capacity, coords = c("lon", "lat"))) %>% 
  st_set_crs("WGS84")

## Plotting to make sure that we have overlaps
# tm_shape(subregion_shapes) +
#   tm_polygons() +
#   tm_shape(plant_points) +
#   tm_dots()

overlap <- st_within(plant_points, subregion_shapes)

## Going through plant_points and get the numeric value that is outputted and store that indexed value into subregion_vector()
subregion_vector <- c()
for (i in seq_along(overlap)) {
  subregion_vector[i] <- subregion_shapes$zip_subregi[overlap[[i]]]
}

## Adding column
plant_capacity <- plant_capacity %>% 
  mutate(subregion = subregion_vector)
```

### Merging data

```{r}
plant_emissions_by_subregion <- left_join(plant_capacity, emissions_by_subregion, by = "subregion")
```

### Importing AISI data

```{r, message=FALSE}
AISI_regions_readin <- readxl::read_excel(here("data", "AISI_regions.xlsx"), sheet = "Regions by State") %>% 
  clean_names()

# AISI_city_readin <- readxl::read_excel(here("data", "AISI_regions.xlsx"), sheet = "Region by City") %>% 
#   clean_names()

AISI_data_readin <- readxl::read_excel(here("data", "AISI_data.xlsx"), sheet = "AISI Production Values") %>% 
  clean_names()
```

NE = Northeast

GL = Great Lakes

MW = Midwest

S = Southern

W = Western

```{r}
AISI_regions <- AISI_regions_readin %>% 
  filter(steel_plant_in_gspt == "yes")

AISI_data <- AISI_data_readin %>% 
  mutate(week_end_date = week_end_date_1
         , NE_util = north_east_region_capacity_utilization
         , GL_util = great_lakes_region_capacity_utilization
         , MW_util = midwest_region_capacity_utilization
         , S_util = southern_region_capacity_utilization
         , W_util = western_region_capacity_utilization) %>% 
  select(week_end_date
         , NE_util
         , GL_util
         , MW_util
         , S_util
         , W_util) %>% 
  filter(week_end_date < "2022-01-01") ## Missing January 2021 data

AISI_data <- AISI_data[-1,]
```

### Merging All Data

```{r}
regional_plant_emissions <- left_join(plant_emissions_by_subregion, AISI_regions, by = "state") %>% 
  select(-c(steel_plant_in_gspt
            , state_abbreviation)) %>% 
  mutate(region = case_when(municipality %in% c("Alton"
                                                , "Sterling"
                                                , "Peoria"
                                                , "Granite City"
                                                , "Mansfield"
                                                , "Middletown") ~ "Midwest"
                            , municipality %in% c("Riverdale"
                                                  , "Chicago"
                                                  , "Bourbonnais"
                                                  , "Cuyahoga Heights"
                                                  , "Cleveland"
                                                  , "Toledo"
                                                  , "Lorain") ~ "Great Lakes"
                            , municipality %in% c("Mingo Junction"
                                                  , "Youngstown"
                                                  , "Canton") ~ "North East"
                            , TRUE ~ as.character(region)))
```

### Emissions Intensity

```{r}
## co2e_tonnes_per_mwh is from eGrid and is pounds of CO2e per MWH of electricity produced per grid location (not regional location)

## Global Efficiency Intelligence states that it takes 710 KWH to produce 1 tonne of steel. "Global Efficiency Intelligence: Industrial Electrification in U.S. States"
## MWH per tonne of steel
eaf_MWH_per_tonne <- 710 / 1000

## emissions_intensity is pounds of CO2e per tonne of steel (the amount of co2e produced for every tonne of steel produced)

AISI_longer <- AISI_data %>% 
  pivot_longer(cols = c(NE_util, GL_util, MW_util, S_util, W_util)
                 , names_to = "region"
                 , values_to = "utilization") %>% 
  mutate(region = case_when(region == "NE_util" ~ "North East"
                   , region == "GL_util" ~ "Great Lakes"
                   , region == "MW_util" ~ "Midwest"
                   , region == "S_util" ~ "Southern"
                   , region == "W_util" ~ "Western"))
```

### Finding Weekly Scope 2 values

```{r}
## eaf_capacity is in thousand tonnes per year and we are standardizing to tonnes per week
## Converting to per thousand tons for weekly co2e
scope2_plant_emissions_long <- left_join(regional_plant_emissions, AISI_longer, by = "region") %>%
  mutate(emissions_intensity_co2e_tonne_per_tonne = (co2e_tonnes_per_mwh * eaf_MWH_per_tonne)
         , eaf_capacity_tonnes_of_steel_per_week = ((eaf_capacity * 1000) / 52)
         , scope2_co2e_tonnes_per_week = (eaf_capacity_tonnes_of_steel_per_week * utilization * emissions_intensity_co2e_tonne_per_tonne)) %>% 
  mutate_if(is.numeric, round, 2)

scope2_plant_emissions_wide <- scope2_plant_emissions_long %>% 
  select(-utilization) %>% 
  pivot_wider(names_from = week_end_date
           , values_from = scope2_co2e_tonnes_per_week)
```

### Finding Yearly Scope 2 values

```{r}
yearly_scope2 <- scope2_plant_emissions_long %>% 
  group_by(plant_id, plant_name, address, lat, lon) %>% 
  summarize(yearly_scope2_co2e_tonnes = sum(scope2_co2e_tonnes_per_week)) %>% 
  ungroup()
```

### Extracting zip codes

```{r}
## Extracting zip code from address
zipcode_extraction <- yearly_scope2 %>% 
  mutate(address2 = address) %>% 
  separate(col = address2, into = c("zip_code_part1", "zip_code_part2", "zip_code_part3", "zip_code_part4", "zip_code_part5"), sep = ",") %>% 
  select(-c(zip_code_part1, zip_code_part5)) %>% 
  mutate(zip_code_part2 = parse_number(zip_code_part2)
         , zip_code_part3 = parse_number(zip_code_part3)
         , zip_code_part4 = parse_number(zip_code_part4)
         , zip_code = coalesce(zip_code_part2, zip_code_part3, zip_code_part4)) %>% 
  select(-c(zip_code_part2, zip_code_part3, zip_code_part4))

checking_unique_zipcodes_scope2 <- zipcode_extraction %>% 
  group_by(zip_code) %>% 
  summarise(count = n()) %>% 
  filter(count < 2) %>% 
  drop_na() %>%
  ungroup() %>% 
  select(zip_code)

unique_scope2_zip_codes <- checking_unique_zipcodes_scope2 %>% 
  left_join(zipcode_extraction, by = "zip_code")
```

### Reading in Greenhouse Gas Emissions data (scope 1)

```{r}
scope1_readin <- readxl::read_excel(here("data", "GHG_flight_scope1.xls"), skip = 5, na = c("", 0)) %>% 
  clean_names()

scope1 <- scope1_readin %>% 
  select(c(zip_code
           , ghg_quantity_metric_tons_co2e
           , ghgrp_id)) %>% 
  drop_na()

checking_unique_zipcodes_scope1 <- scope1 %>% 
  group_by(zip_code) %>% 
  summarise(count = n()) %>% 
  filter(count < 2) %>% 
  ungroup() %>% 
  select(zip_code)

unique_scope1_zipcodes <- checking_unique_zipcodes_scope1 %>% 
  left_join(scope1, by = "zip_code")
```

### Joining scope 1 and scope 2 zip codes

```{r}
## Manually checked zip codes and addresses and plant names and the following are for sure a match
scope_1_and_2_zipcode_matches <- unique_scope1_zipcodes %>% 
  right_join(unique_scope2_zip_codes, by = "zip_code") %>% 
  drop_na() %>% 
  rename(yearly_scope1_co2e_tonnes = ghg_quantity_metric_tons_co2e) %>% 
  relocate(address, .after = zip_code) %>% 
  relocate(plant_name, .after = address)
```

### Anti-Joining to find remaining zip codes

```{r}
remaining_zipcodes <-  zipcode_extraction %>% 
  anti_join(scope_1_and_2_zipcode_matches, by = "zip_code")

## Making sure zip codes match if they are in 
corrected_ghgids <- remaining_zipcodes %>% 
  mutate(ghgrp_id = case_when(plant_id == "SUS00002" ~ "1003268"
                              , plant_id == "SUS00007" ~ "1003668"
                              , plant_id == "SUS00015" ~ "1000394"
                              , plant_id == "SUS00061" ~ "1001699"
                              , plant_id == "SUS00042" ~ "1004616"
                              , plant_id == "SUS00019" ~ "1002977"
                              , plant_id == "SUS00025" ~ "1007642"
                              , plant_id == "SUS00032" ~ "1007921")
         , ghgrp_id = as.numeric(ghgrp_id)
         ) %>% 
  left_join(scope1, by = "ghgrp_id") %>% 
  select(-c(zip_code.x, zip_code.y)) %>% 
  rename(yearly_scope1_co2e_tonnes = ghg_quantity_metric_tons_co2e)
```

### Combining all of our data to get a final dataset with scope 1 and 2 values

```{r}
yearly_scope_1_and_2_emissions <- scope_1_and_2_zipcode_matches %>% 
  full_join(corrected_ghgids) %>% 
  select(-zip_code) %>% 
  relocate(c(plant_id, lat, lon), .before = address) %>% 
  relocate(c(ghgrp_id, plant_name), .after = plant_id) %>% 
  mutate(yearly_co2e_tonnes = yearly_scope1_co2e_tonnes + yearly_scope2_co2e_tonnes, .after = address)
```

## Code that we have tried throughout this document

```{r}
## Importing state electricity values
# egrid2021_data <- readxl::read_excel(here("data", "eGRID2021_data.xlsx"), sheet = "ST21") %>% 
#   clean_names()
```

```{r}
## Reading in egrid data where we were looking at state values
# state_emissions <- egrid2021_data %>% 
#   mutate(co2e = state_annual_co2_equivalent_combustion_output_emission_rate_lb_m_wh
#          , state = state.name[match(state_abbreviation, state.abb)]) %>% 
#   select(c(state, co2e))
# 
# state_emissions <- state_emissions[-1,]
```

##### Attempting to join scope 1 and scope 2 data by different columns and attempts at standardizing both

```{r}
## Trying to standardize address

# gem_data <- gem_data %>% 
#   filter(country == "United States"
#          , status == "operating"
#          , nominal_eaf_steel_capacity_ttpa != "N/A") %>% 
#   mutate(plant_name = plant_name_english
#          , state = subnational_unit_province_state
#          , eaf_capacity = as.numeric(nominal_eaf_steel_capacity_ttpa)
#          , address = tolower(gsub(",.*", "", location_address))
#          , address = gsub("[.].*", "", address)) %>% 
#   select(plant_name
#          , owner
#          , coordinates
#          , country
#          , state
#          , status
#          , start_date
#          , plant_age_years
#          , eaf_capacity
#          , municipality
#          , address)

# scope1 <- scope1_readin %>% 
#   mutate(address = tolower(gsub(",.*", "", reported_address))
#          , address = case_when(address == "2027 e. state hwy.198" ~ "2027 state hwy 198"
#                                     , TRUE ~ as.character(address))
#          , address = gsub("[.].*", "", address)) %>% 
#   select(-c(reporting_year, subparts, ghgrp_id, reported_address))
# 
# 
# scope_1_and_2 <- scope1 %>% 
#   right_join(scope2_plant_emissions_wide, by = "address")
```

```{r}
## Trying to standardize lat and lon (only 4 matched after rounding down to 3 decimal places)

# yearly_scope2 <- scope2_plant_emissions_long %>% 
#   mutate(lat = round(as.numeric(lat), 3)
#          , lon = round(as.numeric(lon), 3)
#          , coordinates = paste0(lat, ",", lon)) %>% 
#   group_by(plant_name, coordinates) %>% 
#   summarise(yearly_scope2_co2e = sum(scope2_co2e_tonnes_per_week))
# 
# scope1_readin <- readxl::read_excel(here("data", "GHG_flight_scope1.xls"), skip = 5, na = c("", 0)) %>% 
#   clean_names()

# scope1 <- scope1_readin %>% 
#   mutate(lat = round(as.numeric(latitude), 3)
#          , lon = round(as.numeric(longitude), 3)
#          , coordinates = paste0(lat, ",", lon)) %>%
#   select(-c(reporting_year, subparts, ghgrp_id, reported_address, latitude, longitude))
# 
# scope_1_and_2 <- scope1 %>% 
#   left_join(yearly_scope2, by = "coordinates")
```
